\chapter{Motivation}
Ever since my first foray into the field of information security I have been fascinated by the ways in which
different threats to an organisation's security arise. From the ways in which data can be obtained by rummaging
through dumpsters where sensitive documents were dumped without being properly destroyed to sophisticated zero-day
exploits used to distribute malware, the topic that particularly caught my interest was the way in which malicious
payloads can be concealed in relatively mundane looking files.

A perfect example of such file was a malicious Microsoft Word document created by the Lazarus Group \acrlong{APT}
distributed to victims in South Korea via spear phishing. The malware itself was hidden within a \acrfull{BMP} file 
that was itself concealed as a \acrfull{PNG} file. This malicious file was extracted to the victim's computer by 
a macro in the macro-enabled Word document, which is also a very interesting part of the infection process. 

This document piqued my interest for a multitude of reasons, chief among which was the interesting mechanism used
to infect the victim's device, which used a quirk of the \acrshort{PNG} and \acrshort{BMP} file formats and a conversion 
function from the \acrfull{WIA} \acrshort{API} in the \acrfull{VBA} programming language used to write macros 
that can be embedded in Word Documents. 

Using this functionality allowed the malicious payload to be concealed under many layers of file formats while also keeping
the extraction process quite simple, almost routine and ceratinly benign looking. The attack itself was also rather 
creative, using an interesting attack vector and custom toolchain typical for this threat actor.

The core of attacks like this one is that since files are all simply a series of bytes in the end -- the
interpretation is governed by how the operating system or program interacting with the file interprets 
those underlying bytes. 

Due to how file formats are defined, some formats are more suitable for attacks than others
and with the multitude of formats supported and used over the decades of computing, it is only inevitable that 
there would be a way to misuse some format in some way -- that being \acrshort{PNG} in this case. 
I believe it is valuable to recreate this attack to shed light on how potential future attacks could be carried out.

Out of interest as well as scientific rigour I attempt to recreate this malware, and analyse its effectiveness,
foregoing the malicious payload and causing no damage in the process. The main questions I seek to answer are the following:
\begin{itemize}
  \item How can a file hide malicious content?
  \item How can a concealed malicious payload be extracted from a file and executed?
  \item How do the previous questions come together to drop the \acrfull{rat} in the analysed document?
  \item Can the analysed document be recreated? Does the exploit still work?
  \item Are common systems still vulnerable?
\end{itemize}

Thus, the primary goal of this work is to recreate the malicious Microsoft Word document along with the image
payload and secondary mocked \acrfull{EXE} payload without the actual \acrlong{rat}. Recreating this malware should help
verify the reproducibility of the attack and its functionality as well as help gain further
insight into how vulnerable current systems are to a similar attack, if at all. Furthermore, this analysis may yield
advice other than the simple adage of not opening macro-enabled documents. Though, of course, this is always the best
protection mechanism against malware using \acrshort{VBA} macros as their attack vector.

To achieve this goal we create a facsimile of the malicious document as well as the payload it carried.
This recreation is based on the postmortem report of the attack written by Hossein Jazi. 
The first part of the recreation is a dummy \acrlong{EXE} containing a simple program that indicates the 
system would have been compromised if the attack was real. This \acrshort{EXE} will be hidden in the \acrshort{PNG}
and embedded in the macro-enabled Word document, analogous to the attack.
Finally, I execute this faux-malicious document inside a virtual machine running the Windows operating system
and track how it executes in comparison to the original attack. 

The metrics for measuring the success of this experiment are rather simple -- recreating the attack in its full
scale is deemed a full success, while failure after at least one part of the attack succeeds is deemed a partial
success. We also keep an eye out on when or whether antivirus software detects the payload.

In summary, recreating this attack can lend insight into how file formats can be misused to carry malicious payloads and
avoid detection while doing so. It also serves as an effort to validate the previous research done of this malware and
make sure the results of that research are reproducible. Furthermore, the alternative setup using a facsimile of the
malicious file may provide additional insight that had previously gone unnoticed.

\clearpage

